GWAS datasets for FDR benchmarking

### 2017-09-13 datasets/GWAS/GWAS.Rmd
add chronotype dataset

*p-value
*effect size
*standard error
*additional covariates (optional)

Note that while knitting this document will download the data, I'm not committing 
the data itself to this repo. See the knitted `.html` document if you do not 
wish to download the datasets but still wish to see the output.

@@

```{bash, GWAS1-download}
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

@@

```{bash, GWAS1-unzip}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
```

@@

```{r, GWAS1-verify-contents}
library(data.table)
bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
head(bmi)
```

@@

- [x] p-value
- [x] effect size
- [x] standard error
- [x] additional covariates: 
  *`N`: Number of samples with this SNP - BMI association measured
  *`Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

>>>>

- p-value
- effect size
- standard error
- additional covariates (optional)

Note that while running all the bash code in this document will download the data, 
some of it doesn't seem to work properly with knitting (e.g. the ftp download 
in the second dataset), and I'm not committing 
the data itself to this repo. Run the bash code in the terminal separately to 
download the data yourself, and see the knitted `.html` document if you do not 
wish to download the datasets but still wish to see the output.

@@

```{bash, GWAS1-download, eval=FALSE}
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

@@

```{bash, GWAS1-unzip, eval=FALSE}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
```

@@

```{r, GWAS1-verify-contents}
library(data.table)
bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
dim(bmi)
head(bmi)
```

@@

- `p_value`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
  - `N`: Number of samples with this SNP - BMI association measured
  - `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

for 2,458,133 SNPs.

```{r, GWAS1-cleanup}
rm(bmi)
```

### GWAS for Self-reported Chronotype

A [study by Hu et al. (Nature Comm, 2016)](https://www.ncbi.nlm.nih.gov/pubmed/?term=26835600) presented a GWAS of 89,283 individuals that identified genetic variants associated with self-reporting of being a morning person.

Let's download the data from the `ebi.ac.uk` ftp site:

```{bash, GWAS2-download, eval=FALSE}
ftp ftp://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/
cd KhorCC_27064256_GCST003467
get "Khor-27064256.txt"
exit
```
Next, we'll read the `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS2-verify-contents}
glauc <- fread("Khor-27064256.txt", header=TRUE)
dim(glauc)
head(glauc)
```

- `pvalue`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
  - `gt.rate`: ?
  - `freq.b`: ?
  - `dose.b`: ?

for 10,000 SNPs - unfortunately, looks like they only give us the *top* 
10K SNPs, so this is not useful for our purposes.

```{r, GWAS2-cleanup}
rm(chrono)
```

### GWAS for Asthma

Let's download the data from the Gabriel site

```{bash, GWAS2-download, eval=FALSE}
curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
```
